---
title: "5-2-aus-hist-max-temp"
output: html_document
---

## Australian historical maximum temperature {#historicaltmax}

```{r}
load(here::here("data/historical_tmax.rda"))
```


The GHCN provides daily climate measures from stations across the world. The data used here (\code{historical_tmax}) is a subset extracted using the package \pkg{rnoaa} [@rnoaa], containing the records of maximum temperature for `r nrow(historical_tmax)` Australian stations from `r range(historical_tmax$first_year)[1]` through `r range(historical_tmax$first_year)[2]` and provides information also on the latitude, longitude and elevation of each of the  stations. <!-- This data  is already cast into a \code{cubble} data object (please see Appendix X for the code to extract  the data from XX and prepare the data into \code{cubble} form), with \code{id} as the key, \code{date} as the index, and \code{c(longitude, latitude)} as the coordinates.--> The goal of this example is to compare the monthly average maximum temperature between two periods, 1971-1975 and 2016-2020, for stations in Victoria and New South Wales (NSW), using a glyph map.

First, the stations need to be filtered to those in Victoria and NSW by using the station identifiers, stored within the 11 digits of the \code{id} variable entries. The country code is in the first 5 digits (Australia is represented by "ASN00") and the next 6 digits encode the station following the Australian Bureau of Meteorology (BOM) [@bom] coding protocols. The NSW stations correspond to entries in the range 46-75 and the Victorian stations to 76-90. Filtering Victoria and NSW stations is a *spatial operation* and hence uses the nested \code{cubble}: 

<!-- BOM link: http://www.bom.gov.au/climate/cdo/about/site-num.shtml -->

```{r echo = TRUE}
tmax <- historical_tmax |>
  filter(between(as.numeric(stringr::str_sub(id, 7, 8)), 46, 90))
```

Next, the monthly maximum average temperature is  calculated for both periods. This is a *temporal operation* requiring a switch into the long \code{cubble} using the \code{face_temporal()} function. In addition, a new indicator for the two time periods of interest is created before the calculation of monthly averages:

```{r echo = TRUE}
tmax <- tmax |>
  face_temporal() |> 
  mutate(month = lubridate::month(date),
         group = as.factor(
           ifelse(lubridate::year(date) > 2015,
           "2016 ~ 2020", "1971 ~ 1975"))) |>
  group_by(month, group) |> 
  summarise(tmax = mean(tmax, na.rm = TRUE))
```

A quick check on the number of observations for each location is made, revealing that there are several with less than 24 observations -- these stations lack temperature values for some months. In this example, those stations are removed by switching to a long \code{cubble} to operate on the spatial component over time, and then, move back into the nested \code{cubble} (to make the glyph map):

```{r echo = TRUE}
tmax |> 
  face_spatial() |> 
  mutate(n = nrow(ts)) |>
  arrange(n) |>
  pull(n) |> 
  head(10)

tmax <- tmax |> 
  face_spatial() |> 
  filter(nrow(ts) == 24) |>
  face_temporal()
```

In order to create a glyph map displaying the monthly series  (Figure \ref{fig:basic-manip}), the spatial variables need to be unfolded with the temporal variables. The reason being that the major (\code{long}, \code{lat}) and minor (\code{month}, \code{tmax}) coordinates need to be on the same table to create the glyph map. The \code{geom_glyph()} function does both the transformation and the plotting. 

```{r eval = FALSE, echo = TRUE}
nsw_vic <- ozmaps::abs_ste |>
  filter(NAME %in% c("Victoria", "New South Wales"))

ggplot() +
  geom_sf(data = nsw_vic,
          fill = "transparent", color = "grey",
          linetype = "dotted") +
  geom_glyph(data = tmax,
             aes(x_major = long, x_minor = month,
                 y_major = lat, y_minor = tmax,
                 group = interaction(id, group), color = group),
             width = 1, height = 0.5) +
  ...
```

```{r}
cobar <- tmax |> filter(id == "ASN00048027") |>
  ggplot(aes(x = month, y = tmax, color = group)) +
  geom_line() +
  scale_color_brewer(palette = "Dark2", guide = "none") +
  scale_x_continuous(
    breaks = seq(1, 12, 1), 
    labels = c("J", "F", "M", "A",
               "M", "J", "J", "A",
               "S", "O", "N", "D")
    ) +
  labs(x = "", y  = "Temp (C)") +
  theme_bw() + 
  theme(aspect.ratio = 0.3)

```


```{r basic-manip, out.width="100%", fig.height = 5, fig.width = 10, fig.cap="A glyph map of the monthly maximum average temperature for weather stations in  Victoria and New South Wales (NSW) for the periods (1971-1975, 2016-2020). The corresponding average time series for the cobar station are display on the top left corner. From the glyph map we can observe that the monthly trend is similar for all locations (low in the winter, high in the summer), and small increased temperatures, particularly in late summer can be seen at most stations in NSW."}
tmax <- tmax |> unfold(long, lat)
box_df <- tmax |> face_spatial() |> filter(id == "ASN00048027")
nsw_vic <- ozmaps::abs_ste |> filter(NAME %in% c("Victoria","New South Wales"))

p1 <- ggplot() + 
  geom_sf(data = nsw_vic, fill = "grey95", 
          color = "white") + 
  geom_glyph(
    data = tmax, 
    aes(x_major = long, x_minor = month, 
        y_major = lat, y_minor = tmax,
        group = interaction(id, group), color = group),
    width = 1.2, height = 0.8) +
    geom_rect(
      data = box_df,
      aes(xmin = long - 0.6, xmax = long + 0.6,
          ymin = lat - 0.12, ymax = lat + 0.35),
      fill = "transparent", color = "black") +
  scale_color_brewer("", palette = "Dark2") + 
  coord_sf(xlim = c(141, 154), ylim = c(-39, -28.5)) + 
  ggthemes::theme_map() +
  theme(legend.position = "bottom", legend.text = element_text(size = 10)) + 
  guides(color = guide_legend(override.aes = list(size=2)))

(p1 | cobar)  + 
  patchwork::plot_layout(width = c(1,1)) + 
  plot_annotation(tag_levels = 'a')

```

Glyph maps work well to explore temporal patterns across spatial locations, particularly when the spatial locations are gridded. In this example, they are irregularly spaced, which can result in overlapping glyphs obscuring each other. To fix this, one could aggregate data from nearby stations. An example of this is included in the Appendix. 
